---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

### Indroduction  
I am interested in studying data on employment and productivity, because I am interested to see what kind of correlations there could be between things such as hours worked, compensation, profits, and productivity. I think it will be useful to about any relationships between these factors, so that we can figure out the best way to have people work.  

### Dataset 
Before we start we will need to load a library called "tidyverse", since we will be using it later on.
```{r}
suppressMessages(library("tidyverse"))
```

Now lets load in the data from "https://www.bls.gov/lpc/major_sector_lpc_data.csv".
```{r}
data <- read.csv("https://www.bls.gov/lpc/major_sector_lpc_data.csv")
```

Now let's take a look at our data to see what data there is.
```{r}
head(data)
nrow(data)
ncol(data)
```

Now that we have an idea of what the data looks like, lets fix up the column names a little so that it is easier to read. We will change all the spots where it says "2012...100" to "2012.100" and we will also change all the spots where there's two periods to only have one.
```{r}
names(data) <- gsub(x = names(data), pattern = "\\.\\.\\.", replacement = "\\.")
names(data) <- gsub(x = names(data), pattern = "\\.\\.", replacement = "\\.")
```

### Source  
This dataset comes from the U.S. Bureau of Labor Statistics which is a very reliable source since they are part of the U.S. government. They have documented how and where they got the data from, so you can know that it is reliable.  

### Variables
Now we will take a look at all the variables in the dataset. We will first print them all out, then look at each one individually.
```{r}
colnames(data)
```

####The Categorical Variables
**Major.Sectors:** The type of sector that the observation was made about.  
**Year:** The year that the observation was made.  
**Quarter:** The quarter that the observation was made.  
  
####For the continuous variables there's three different variations:
  * **2012.100:** which is where the values are scaled so that at the year 2012 the value will be 100.
  * **percent.change.from.same.quarter.a.year.ago:** which is the percent change from that quarter from the year before.
  * **percent.change.from.previous.quarter.at.annual.rate:** which is the percent change from the previous quarter at an annual rate.
  
####The actual variables are:
**Employment:** The total numbers of wage and salary workers, unincorporated self-employed workers, and unpaid family workers working within business establishments. An individual who works multiple jobs at separate establishments would have each job included in the number of employees.  
**Average.weekly.hours.worked:** The total number of hours worked over a specified period of time, divided by the total number of weeks worked in the time period.   
**Hours.worked:** The total number of hours worked by wage and salary workers, unincorporated self-employed workers, and unpaid family workers to produce output.  
**Output:** The amount of goods and services produced.    
**Current.dollar.output:** The value of goods and services produced, expressed in current dollars.   
**Compensation:** Monetary and non-monetary payments to or on behalf of individuals for labor services used to produce output. Compensation may include salaries, wages, bonuses, contributions to benefits plans, and other forms of payment.  
**Nonlabor.payments:** The value of output less the cost for labor services.    
**Labor.productivity:** The efficiency at which labor hours are utilized in producing output of goods and services, measured as output per hour of labor.   
**Hourly.compensation:** The sum of wage and salary accruals and supplements to wages and salaries per hour of labor services used to produce output. Wage and salary accruals consist of the monetary remuneration of employees. Supplements to wages and salaries consist of employer contributions for social insurance and employer payments (including payments in kind) to private pension and profit-sharing plans, group health and life insurance plans, privately administered workers' compensation plans.  
**Unit.labor.costs:** The compensation for labor services used to produce each unit of goods and services.    
**Unit.nonlabor.costs:** The costs (payments) for each unit of goods and services produced after excluding labor costs and profits from the cost (value) of output.  
**Unit.nonlabor.payments:** The costs (payments) for each unit of goods and services produced after excluding labor costs and profits from the cost (value) of output.    
**Implicit.price.deflator:** An index of the change over time in the value of output relative to a base period.  
**Real.hourly.compensation:** The average monetary payments made to individuals for an hour of labor service, adjusted for inflation.  
**Output.per.job:** Output per job represents the amount of goods or services a person engaged in a particular occupation (job) can produce over an interval of time, such as a year, regardless of the actual number of hours worked. Output per job can be thought of a measure of output per person that is adjusted for the notion that a person can hold multiple jobs at the same time.  
**Labor.share:** The portion of the total costs to produce output that can be attributed to the cost of labor.  
**Profits:** The portion of total income from the sale of goods and services that exceeds the total costs of producing those goods and services.  
**Unit.profits:** Corporate profits before taxes with inventory valuation and capital consumption adjustments. Total unit costs, less labor costs, are profits.  
**Total.unit.costs:** The total payments (costs) for each unit of output produced where costs are measured as a difference between total income and profits.  

###Seperate Data into Different Tables
Now let's split this data into a few different tables so that each table will have one observational unit. This means that each table will be more focused, which will make using the data easier.    
   
The first table will be called from_year_ago and contain all the variables that are measuring the percent change from the same quarter a year ago.
```{r}
from_year_ago <- tibble(major_sectors=data$Major.Sectors, year=data$Year, quarter=data$Quarter,
                        Employment=data$Employment.percent.change.from.same.quarter.a.year.ago.,
                        Average_weekly_hours_worked=data$Average.weekly.hours.worked.percent.change.from.same.quarter.a.year.ago.,
                        Hours_worked=data$Hours.worked.percent.change.from.same.quarter.a.year.ago.,
                        Output=data$Output.percent.change.from.same.quarter.a.year.ago.,
                        Current_dollar_output=data$Current.dollar.output.percent.change.from.same.quarter.a.year.ago.,
                        Compensation=data$Compensation.percent.change.from.same.quarter.a.year.ago.,
                        Nonlabor_payments=data$Nonlabor.payments.percent.change.from.same.quarter.a.year.ago.,
                        Labor_productivity=data$Labor.productivity.percent.change.from.same.quarter.a.year.ago.,
                        Hourly_compensation=data$Hourly.compensation.percent.change.from.same.quarter.a.year.ago.,
                        Unit_labor_costs=data$Unit.labor.costs.percent.change.from.same.quarter.a.year.ago.,
                        Unit_nonlabor_costs=data$Unit.nonlabor.costs.percent.change.from.same.quarter.a.year.ago.,
                        Unit_nonlabor_payments=data$Unit.nonlabor.payments.percent.change.from.same.quarter.a.year.ago.,
                        Implicit_price_deflator=data$Implicit.price.deflator.percent.change.from.same.quarter.a.year.ago.,
                        Real_hourly_compensation=data$Real.hourly.compensation.percent.change.from.same.quarter.a.year.ago.,
                        Output_per_job=data$Output.per.job.percent.change.from.same.quarter.a.year.ago.,
                        Labor_share=data$Labor.share.percent.change.from.same.quarter.a.year.ago.,
                        Profits=data$Profits.percent.change.from.same.quarter.a.year.ago.,
                        Unit_profits=data$Unit.profits.percent.change.from.same.quarter.a.year.ago.,
                        Total_unit_costs=data$Total.unit.costs.percent.change.from.same.quarter.a.year.ago.)
```

The second table will be called from_previous_quarter and contain all the variables that are measuring the percent change from the previous quarter at an annual rate.
```{r}
from_previous_quarter <- tibble(major_sectors=data$Major.Sectors, year=data$Year, quarter=data$Quarter,
                        Employment=data$Employment.percent.change.from.previous.quarter.at.annual.rate.,
                        Average_weekly_hours_worked=data$Average.weekly.hours.worked.percent.change.from.previous.quarter.at.annual.rate.,
                        Hours_worked=data$Hours.worked.percent.change.from.previous.quarter.at.annual.rate.,
                        Output=data$Output.percent.change.from.previous.quarter.at.annual.rate.,
                        Current_dollar_output=data$Current.dollar.output.percent.change.from.previous.quarter.at.annual.rate.,
                        Compensation=data$Compensation.percent.change.from.previous.quarter.at.annual.rate.,
                        Nonlabor_payments=data$Nonlabor.payments.percent.change.from.previous.quarter.at.annual.rate.,
                        Labor_productivity=data$Labor.productivity.percent.change.from.previous.quarter.at.annual.rate.,
                        Hourly_compensation=data$Hourly.compensation.percent.change.from.previous.quarter.at.annual.rate.,
                        Unit_labor_costs=data$Unit.labor.costs.percent.change.from.previous.quarter.at.annual.rate.,
                        Unit_nonlabor_costs=data$Unit.nonlabor.costs.percent.change.from.previous.quarter.at.annual.rate.,
                        Unit_nonlabor_payments=data$Unit.nonlabor.payments.percent.change.from.previous.quarter.at.annual.rate.,
                        Implicit_price_deflator=data$Implicit.price.deflator.percent.change.from.previous.quarter.at.annual.rate.,
                        Real_hourly_compensation=data$Real.hourly.compensation.percent.change.from.previous.quarter.at.annual.rate.,
                        Output_per_job=data$Output.per.job.percent.change.from.previous.quarter.at.annual.rate.,
                        Labor_share=data$Labor.share.percent.change.from.previous.quarter.at.annual.rate.,
                        Profits=data$Profits.percent.change.from.previous.quarter.at.annual.rate.,
                        Unit_profits=data$Unit.profits.percent.change.from.previous.quarter.at.annual.rate.,
                        Total_unit_costs=data$Total.unit.costs.percent.change.from.previous.quarter.at.annual.rate.)
```

The third table will be called where_2012_is_100 and contain all the variables that are scaled so that at the year 2012 the value will be 100.
```{r}
where_2012_is_100 <- tibble(major_sectors=data$Major.Sectors, year=data$Year, quarter=data$Quarter,
                        Employment=data$Employment.2012.100.,
                        Average_weekly_hours_worked=data$Average.weekly.hours.worked.2012.100.,
                        Hours_worked=data$Hours.worked.2012.100.,
                        Output=data$Output.2012.100.,
                        Current_dollar_output=data$Current.dollar.output.2012.100.,
                        Compensation=data$Compensation.2012.100.,
                        Nonlabor_payments=data$Nonlabor.payments.2012.100.,
                        Labor_productivity=data$Labor.productivity.2012.100.,
                        Hourly_compensation=data$Hourly.compensation.2012.100.,
                        Unit_labor_costs=data$Unit.labor.costs.2012.100.,
                        Unit_nonlabor_costs=data$Unit.nonlabor.costs.2012.100.,
                        Unit_nonlabor_payments=data$Unit.nonlabor.payments.2012.100.,
                        Implicit_price_deflator=data$Implicit.price.deflator.2012.100.,
                        Real_hourly_compensation=data$Real.hourly.compensation.2012.100.,
                        Output_per_job=data$Output.per.job.2012.100.,
                        Labor_share=data$Labor.share.2012.100.,
                        Profits=data$Profits.2012.100.,
                        Unit_profits=data$Unit.profits.2012.100.,
                        Total_unit_costs=data$Total.unit.costs.2012.100.)
```

###Summarizing the Data
####Lets take a look at some descriptive statistics. We will use the summary function to get information about the variables in the various tables. 

First lets look at the table measuring the percent change from the same quarter a year ago.
```{r}
summary(from_year_ago)
```
  

Next, we'll look at the table measuring the percent change from the previous quarter at an annual rate.
```{r}
summary(from_previous_quarter)
```
  

Finally, we'll look at the table where the values are scaled so that at the year 2012 the value will be 100.
```{r}
summary(where_2012_is_100)
```


####Now lets look at some visualizations of the data.

First we can look at the percent change in Employment from the same quarter a year ago, and have different lines based on the sector. Some interesting things we can see from this is that there was a big increase in empoyment in the nonfinancial corporations sector around 1950, and there was a big decrease in the employment in manufacturing a little before 2010.
```{r}
ggplot(from_previous_quarter, aes(x=year, y=Employment, group=major_sectors)) + geom_line(aes(color=major_sectors)) + labs(title="Percent Change in Employment from the Same Quarter a Year Ago", x="Year", y="Percent Change in Employment", color = "Major Sectors")
```

Next we can look at percent change in hourly compensation from the previous quarter at an annual rate, and have different lines based on the sector. Some interesting things we can see from this graph is that it was increasing a lot for all sectors from about 1965 to 1980, then it started to increase much slower.
```{r}
ggplot(from_year_ago, aes(x=year, y=Hourly_compensation, group=major_sectors)) + geom_line(aes(color=major_sectors)) + labs(title="Percent Change in Hourly Compensation from the Previous Quarter at an Annual Rate", x="Year", y="Percent Change in Hourly Compensation", color = "Major Sectors")+ theme(plot.title = element_text(size=12))
```

Finally we will look at labor productivity where it's scaled so that at the year 2012 the value will be 100, and have different lines based on the sector. An interesting thing that we can see from this is that manufacturing started out much lower that the other sectors, then increased quickly until around 2010 then slowed down and even started to decrease. 
```{r}
ggplot(where_2012_is_100, aes(x=year, y=Labor_productivity, group=major_sectors)) + geom_smooth(aes(color=major_sectors)) + labs(title="Labor Productivity", x="Year", y="Labor Productivity", color = "Major Sectors")
```

###Research Questions
#####There are many potential research questions that could be answered by this data, but here are some that I think could be interesting:  
  * How does increased productivity affect other factors such as employment, hours worked, and compensation?
    + This could have interesting implications because we could see if increasing productivity is good for empoylees or not.
  * What kind of effect does average weekly hours worked have on compensation?
    + It would be interesting to see what kind of relationship there is (linear or other), to be able to tell if it is worth it to work more hours.